Icicle: write once, run once

### 2016-05-30 2016fhpc/section/01-Introduction.tex
Shift cut out text to separate files
Copy edits to intro

Joint queries such as the stocks example can be evaluated in a streaming, incremental fashion, which allows us to update the result as we receive new data. As a counter example, suppose we have a table with two fields @key@ and @value@, and we wish to find the mean of values whose key matches the last one in the table. We might try something like:
\begin{code}
  table kvs { key : Date; value : Real }
  query avg = let k = last key
              in  filter (key == k) of mean value;
\end{code}
Unfortunately, although the \emph{result} we desire is computable, the \emph{algorithm} implied by the above query cannot be evaluated in an inceremental fashion. While streaming through the table we always have access to the last key in the stream, but to perform the filter to select rows that match this key would require that we stream through the table again from the start. We need a better solution. With this in mind, the contributions of this paper are:
\begin{itemize}
\item
We present a domain specific language and type system that guarantees any two queries on the same input table can be fused, and allows queries to be resumed as new data is received\sref{s:IcicleSource};

@@

\item
We present production benchmarks of Icicle compiled code on our smaller single-disk datasets, which compare favourably against standard unix utilities, and outperform an existing R feature generation system by several orders of magnitude\sref{s:Benchmarks}.
\end{itemize}

@@

% This is achieved by having a very restricted query language with no sorting, and allowing only a single-pass over the data.
% \ben{Maybe leave mention of sorting until later sections}.

% \ben{Rewrite the examples using this syntax?}
% In Icicle, the above queries are expressed as:
% \begin{code}
% table stocks { open : Int; close : Int; }
% query more = filter open > close in count;
% query less = filter open < close in count;
% query mean = filter open > close in sum open / count;
% \end{code}

% \ben{the abstract already says this.}
% For queries over a large amount of data, care must be taken to reduce the number of iterations over the input data. When performing multiple queries, it is important to limit the amount of duplicate work and iterations.

% \ben{Start with the example.}
% \begin{code}
%   SELECT COUNT(*) FROM stocks WHERE open > close;
%   SELECT COUNT(*) FROM stocks WHERE open < close;
% \end{code}
% As both queries are over the same input data set, we would like to evaluate them using a single pass over this data. As they use different @WHERE@ clause filters, we must fuse them by lifting the filter predicate into the expression being evaluated:

% In order to fuse these so that both are computed in a single query, some ingenuity is required. This is because the @WHERE@ clause filters over the entire query, but when putting both in one query we need two different filters.

% \begin{code}
%   SELECT COUNT(IF(open > close, 1, 0))
%        , COUNT(IF(open < close, 1, 0))
%   FROM stocks;
% \end{code}
% Suppose we also want to compute the mean of open price on those days:
% \begin{code}
%   SELECT SUM(open) / COUNT(*)
%   FROM   stocks
%   WHERE  open > close;
% \end{code}

% These queries are then fused together, common subexpression elimination is performed, and efficient C code is generated.

>>>>

Joint queries such as the stocks example can be evaluated in a streaming, incremental fashion, which allows the result to be updated as we receive new data. As a counter example, suppose we have a table with two fields @key@ and @value@, and we wish to find the mean of values whose key matches the last one in the table. We might try something like:
\begin{code}
  table kvs { key : Date; value : Real }
  query avg = let k = last key
              in  filter (key == k) of mean value;
\end{code}
Unfortunately, although the \emph{result} we desire is computable, the \emph{algorithm} implied by the above query cannot be evaluated in an inceremental fashion. While streaming through the table we always have access to the last key in the stream, but to select rows that match this key would require that we stream through the table again from the start. We need a better solution. With this in mind, the contributions of this paper are:
\begin{itemize}
\item
We present a domain specific language that guarantees any set of queries on a shared input table can be fused, and allows the query results to be updated as new data is received\sref{s:IcicleSource};

@@

\item
We present production benchmarks of Icicle compiled code which compare favourably to standard unix utilities, and outperform an existing R feature generation system by several orders of magnitude\sref{s:Benchmarks}.
\end{itemize}

% on our smaller single-disk datasets
% BEN: for streaming applications the absolute size of the input size doesn't affect the performance, so we don't need to apologise.

@@
